\documentclass{notes}
% LITERATURE
\addbibresource{sources.bib}
% COMMANDS
\usepackage{commands}
% THEOREMS
\usepackage{theorems}
% COLOR CODE
\usepackage{ourcolors}

\def\C{\mathbf{C}}
\def\Set{\mathbf{Set}}
\def\Profin{\mathbf{ProFin}}
\def\Lightprofin{\mathbf{Pro_{\mathbb{N}}Fin}}
\def\Disccomp{\mathbf{DiscComp}}

\def\id{\textup{id}}
\def\op{\textup{op}}

\begin{document}

\title{Notes on Condensed Mathematics} 
\author{Kai Cieliebak, Milan Zerbin, ...}
\date{\today}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%% Abstract %%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{abstract} 
%\end{abstract}

\maketitle

%\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Light condensed sets}\label{sec:light-condensed-sets}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The ordinary text in this section is a transcript of Scholze's Lecture
2 in~\cite{videos}, whose video is available under
\url{https://www.youtube.com/playlist?list=PLx5f8IelFRgGmu6gmL-Kf$\_$Rl$\_$6Mm7juZO}

Text in colours such as {\kai this one} contains details and
explanations from our side.
Current collaboratours with defined colors:
{\kai Kai}, % Red
{\milan Milan}, % Fuchsia
{\filip Filip}, % Blue
{\sadrian Adrian A}, % Dark Cyan
{\adrian Adrian W}, % Green
{\lukas Lukas},
{\luca Luca}, % Purple
\dots

Other relevant references are the lecture
notes~\cite{analytic,condensed,complex}. 

%%%
\subsection{Light profinite sets}
%%%


\textcolor{black}{}
{\milan 
There are two different notions of highly disconnected spaces.
\begin{definition}
    A topological space $X$ is called \emph{totally disconnected}, if any subset of $X$ containing more than one element is disconnected.
    It is called \emph{totally separated}, if any two points $x,y\in X$ can be separated by a clopen (closed and open) $U\subset X$ with $x\in U$ and $y\not\in U$.
\end{definition}
Any totally separated space is totally disconnected:
Let $A\subset X$ be a subset containing two distinct elements $x,y \in A$. Take $U$ a separating clopen set with $x\in U$, $y\not\in U$. 
The complement $U^c = X\setminus U$ of $U$ is clopen aswell.
Then $A = (A\cap U) \cup (A\cap U^c)$ is a decomposition of $A$ into two nonempty open sets, hence $A$ is disconnected.
The converse however is generally not true
\begin{example}
    Consider the space $X = \N\cup \{\infty, \infty^\prime\}$ with discrete topology on $\N$ and open sets around $\infty$ or $\infty^\prime$ containing all except finitely many natural numbers.
    The map $n\mapsto \frac{1}{n}$ identifies this space with the subspace $\{1/n \:\vert\: n\in\N\}\cup\{0,0^\prime\}$ of the \emph{line with two origins} $\R_0^+\times\{1\} \amalg \R_0^+\times\{-1\} / \sim$ where $(x,1)\sim (x,-1)$ whenever $x>0$.

    $X$ is not hausdorff, as the sequence $(n)$ converges both to $\infty$ and $\infty^\prime$.
    The clopen set $\{\infty,\infty^\prime \}\cup \{m \:\vert\: m> n\}$ separates $\infty$ or $\infty^\prime$ from $n$, so all sets containing both an element from $\N$ and one from $\{\infty,\infty^\prime\}$ are disconnected.
    Furthermore the subset $\{\infty,\infty^\prime\}$ inherits the discrete topology, as there are open sets around $\infty$ not intersecting $\{\infty^\prime\}$ in $X$.
    Thus also the subspace $\{\infty^\prime,\infty\}$ is disconnected, hence $X$ is totally disconnected.

    $X$ is however not totally separated: Any open set around $\infty$ contains infinitely many $n\in \N$, its closure then contains $\infty^\prime$ aswell. 
    Hence $\infty$ and $\infty^\prime$ may not be separated by a clopen set.
\end{example}
\begin{lemma}
    A compact hausdorff space is totally disconnected if and only if it is totally separated
\end{lemma}
\begin{proof}
By above, totally separated always implies totally disconnected. 
For the other direction, let $X$ be totally separated and $x,y\in X$. 
Pick open neighbourhoods $x\in U$ and $y\in V$ with $U\cap V = \emptyset$ by the Hausdorff property of $X$.
    Since $X$ is totally disconnected, $U\cap V$ is disconnected and hence admits a clopen decomposition $\widetilde{U}\cup \widetilde{V}$ with $\widetilde{U} \cap \widetilde{V} = \emptyset$.
\end{proof}
}

{\adrian
\paragraph{One-point compactification of $\mathbb N$}

The usual way of defining the one-point compactification of $\mathbb N$ (as a discrete space) to $\mathbb N^* = \mathbb N \cup \{\infty\}$ is by having the previously open sets of $\mathbb N$ (all of them) be open in $\mathbb N^*$, and additionally all those sets containing $\infty$, whose complement is closed and compact inside $\mathbb N$ (equivalently, finite).

To understand $\mathbb N^*$ as a pro-finite set with the corresponding topology (inverse limit-topology), we consider the projective system of finite sets
  \[
    X_n \coloneq \{ k \in \mathbb N \mid k \leq n\} \cup \{ \infty \} = \{0, 1, \dots, n, \infty\}
  \]
  with maps $p^{n}_m \colon X_{n} \longrightarrow X_{m}$ for $n \geq m$ given by
  \[
    k \longmapsto
    \left[
    \begin{array}{lll}
      k
      & ,
      & k \leq m
      \\  \infty
      & ,
      & \textup{$k > m$, esp. if $k = \infty$}
    \end{array}
    \right.
  \]
  These maps commute, in the sense that $p^m_l \circ p^n_m = p^n_l$, which is a necessity for defining the (projective/inverse) limit of this system
  \begin{equation}
    \label{def:compactified_IN}
    \mathbb N^* \coloneq \varprojlim_{n} X_n.
  \end{equation}
  We take this limit in the category of topological spaces, where the sets $X_n$ are discrete spaces.


\paragraph{The underlying set of $\mathbb N^*$:}
Let's first check that this gives the correct set $\mathbb N \cup \{\infty\}$, by constructing a bijection.
An element of $\mathbb N^*$ can be understood as an element of the product $x = (x_n) \in \prod_n X_n$ that fulfills the compatibility conditions
  \[
    p^n_m(x_n) = x_m
  \]
  for all $m \leq n$.
Since $(X_n)$ is a linear system, it suffices to check this for the maps $p^{n+1}_n$, because all the others are compositions of those maps between ``adjacent'' sets:
  \[
    p^n_m = p^{m+1}_m \circ p^{m+2}_{m+1} \circ \cdots \circ p^n_{n-1}.
  \]
  Let's say that we have $x_n = k \neq \infty$ for some $n$.
Then on the one hand we can conclude for all smaller indices $m \leq n$ that
  \[
    x_m = p^n_m(x_n) = p^n_m(k) =
    \left[
    \begin{array}{lll}
      k
      & ,
      & k \leq m
      \\  \infty
      & ,
      & \textup{$k > m$, esp. if  $k = \infty$}
    \end{array}
    \right.,
  \]
  For all bigger indices $m \geq n$ we have $x_m = k$, since $k \neq \infty$ and therefore its only preimage under $p^m_n$ is $k \in X_m$, thus
  \[
    x = (\infty, \dots, \infty, k, k, k, \dots),
  \]
  where $x_k$ is the first finite entry.
This will be the image of $k \in \mathbb N$ under $\mathbb N \cup \{\infty\} \longrightarrow \mathbb N^*$.

The only other $x \in \mathbb N^*$ will be the one that has no finite entry, so
  \[
    x = (\infty, \infty, \dots),
  \]
  which will be the image of $\infty$, which completes the bijection.

Another thing to mention is that there are natural projections
  \[
    p_n \colon \mathbb N^* \longrightarrow X_n, x \longmapsto x_n.
  \]


  \paragraph{Topology on $\mathbb N^*$:}

  The topology on $\mathbb N^*$, viewed as an inverse limit (in the category of topological spaces), is by definition given by the coarsest topology such that these projection maps are continuous.
  (It can also be infered by restriction of the product topology on $\prod_n X_n$, but strictly speaking, this toplogy arises from the same definition, where the product is viewed as an inverse limit.)

  Since the sets $X_n$ are discrete and finite spaces, every open subsets of $\mathbb N^*$ is then a union of the preimages of single points:
  \[
      {p_n}^{-1}(k) = \{ x \mid x_n = k \} =
    \left[
    \begin{array}{lll}
      \{ k \}
      & ,
      & k \leq \infty
      \\  \{ k, k+1, \dots, \infty\}
      & ,
      & k = \infty
    \end{array}
    \right.
  \]
  Restricted to the subspace $\mathbb N$ this gives the discrete topology as expected, which means that all sets not containing $\infty$ are open, whereas sets containing $\infty$ are open if and only if they also contain every natural number bigger than some number $k \in \mathbb N$, i.e.\ their complement is closed and compact (finite) in $\mathbb N$.
This is exactly the same topology that we had put on the set $\mathbb N \cup \{\infty\}$ as the one-point compactification of $\mathbb N$, and the identification described above is therefore a homeomorphism.

}\label{subsec:light-profinite-sets}


{\filip
\begin{example} $\lim_n \Z/p^n \Z = \Z_p.$ 

In order to talk about the (inverse) limit, we need to specify the maps $$p_{ij}: \Z/p^j \Z \to \Z p^i \Z.$$ They are given by $p_{ij}([q]_{p^j}) = [q]_{p^{i}}$ for $i \leq j$. The result of the limit is the set $\Z_p$ of $p$-adic integers. $\Z_p$  is the completion of $\Z$ with respect to the $p$-adic norm $\|n\|_p = \frac{1}{p^k}$, where $n=p^k q $ and $\mathrm{gcd}(p,q)=1$. 

Given a sequence $a_n \in \{0, 1, ..., p\}$, let us show that $S_k= \sum_{i=0}^k a_i p^i$ is a Cauchy sequence in $\| \cdot \|_p$. Indeed, assume $m<n$ 
$$
\|S_n- S_m\|_p = \|\sum_{i=m+1}^n a_i p^i\|_p \leq \frac{1}{p^{m+1}}.
$$
So the limit of $S_n$ exists in $\Z_p$, and we can write $\sum_{n=0}^{+\infty} a_n p^n \in \Z_p$. One can actually see that 
$$
\Z_p= \left\{\sum_{n=0}^{+\infty} a_n p^n \mid a_n \in \{0, ..., p-1\}\right \}.
$$
Now, recall that in the category of topological spaces, the limit $\lim_{i \in I} X_i = X_{\infty}$ is given by
$$
X_{\infty} = \left\{ (x_i)_{i \in I} \in \prod_{i \in I} X_i \mid  p_{ij}(x_j) = x_i \right\}.
$$
Now, each element $q \in \Z / p^j\Z$ has the unique representative of the form $q= \sum_{k=0}^{j-1} a_k p^k$, and 
$$
p_{ij}\left(\sum_{k=0}^{j-1} a_k p^k\right) = \sum_{k=0}^{i-1}a_k p^k.
$$
Hence, for $q = (q_n)_{n \in \N} \in \lim_n \Z /p^n \Z$, and since $p_{ij}(q_j)= q_i$ we have
$$
q_n = \sum_{k=0}^{n-1} a_k p^k.
$$
Note that the coefficients $a_k$ do not depend on $n$. So, we have an obvious bijection between $\Z_p$ and $\lim_n \Z/p^n \Z$ given by
$$
\Z_p \ni \sum_{n=0}^{\infty} a_k p^k \mapsto \left(\sum_{k=0}^{n-1} a_k p^k \right)_n \in \lim_n \Z/p^n \Z.
$$
Now, we need to check that the topologies agree. The topology on the direct limit is the relative topology (as a subset) of the product topology. Since we equip $\Z/ p^k \Z$ with the discrete topology, the basis for the topology on $\lim_n \Z/p^n \Z$ is given by prescribing finitely many values. On the other hand, open balls in $\Z_p$ are precisely the sets where the first $k$ entries are equal to the entries of the center of the ball, and all other entries are arbitrary.
\end{example}
}

{\lukas
We first discuss classical Stone duality between totally disconnected compact Hausdorff spaces and boolean algebras.

\begin{definition}
  A \emph{Stone space} is a totally disconnected compact Hausdorff space.
  We denote the category of Stone spaces and continuous maps by \(\Stone\).
\end{definition}

\begin{definition}
  A \emph{boolean algebra} is a commutative ring \(A\) (with unit \(1\)) such that every element \(a\in A\) is \emph{idempotent}, that is, \(a^2=a\).
  We denote the category of Boolean algebras and ring homomorphisms by \(\Bool\).
\end{definition}

\begin{example}
  The two element set \(\{0,1\}\) becomes a Stone space by equipping it with the discrete topology.
  Moreover, it carries the structure of a boolean algebra by identifying it with \(\F_2\), the finite field with two elements.
  (In fact, \(\F_2\) is a \emph{boolean algebra object} in the category of Stone spaces.)
\end{example}

\begin{example}
  For every Stone space \(X\), the set \(\Cont(X,\F_2)\) of continuous maps from \(X\) to \(\F_2\) becomes a boolean algebra with respect to pointwise addition and multiplication.
  Note that \(\Cont(X,\F_2)\) is in bijection to the set of clopen subsets of \(X\).
  (This is an isomorphism of boolean lattices?)
\end{example}

\begin{lemma}
	\label{lem:stone-spaces}
  \leavevmode
  \begin{enumerate}
    \item
      Arbitrary products of Stone spaces are Stone spaces.
    \item
      Closed subspaces of Stone spaces are Stone spaces.
  \end{enumerate}
\end{lemma}
\begin{proof}
  By Tychonoff, arbitrary products of compact spaces are compact.
  The same holds for Hausdorff spaces and totally disconnected spaces (?).
  \dots
\end{proof}

\begin{example}
  Let \(A\) be a boolean algebra.
  The set \(\Hom(A,\F_2)\) of ring homomorphisms from \(A\) to \(\F_2\) is a closed subset of \(\prod_{a\in A}\F_2\) equipped with the product topology.
  Endowing it with the induced subspace topology, it becomes a Stone space which is called the \emph{spectrum} of \(A\) and denoted by \(\Spec(A)\).

  Missing: Show that this is homeomorphic to the Zariski spectrum!
  Under this homeomorphism, the standard open subsets \(D(a)\), where \(a\in A\), correspond to open sets of the form
  \[\{\phi\in\Spec(A) \mid \phi(a)=1\},\]
  which are henceforth denoted by \(D(a)\) as well.
  In fact, \(D(a)\) is also a closed set, since \(\Spec A\setminus D(a)=D(1-a)\).
  Consequently, the spectrum of a boolean algebra possesses a basis of clopen subsets, which gives another proof that it is totally disconnected.
\end{example}

\begin{thm}[Stone Duality I]
  Assigning to a Stone space \(X\) the boolean algebra \(\Cont(X,\F_2)\), and to a boolean algebra \(A\) its spectrum \(\Spec(A)\), constitutes an equivalence of categories
  \[
    \Stone \simeq \Bool^{\rm op}.
  \]
\end{thm}
\begin{proof}
  First of all, the assignments \(X\to\Cont(X,\F_2)\) and \(A\mapsto\Spec(A)=\Hom(A,\F_2)\) are obviously (contravariantly) functorial (by precomposition).
  We are then to show that they are mutually (pseudo)inverse to each other, that is, we have to find natural isomorphisms \(X\cong\Spec(\Cont(X,\F_2))\) and \(A\cong\Cont(\Spec A,\F_2)\).
  Both of these are defined as evaluation maps.

  More precisely, let us first consider the map
  \[e:X \longrightarrow \Spec(\Cont(X,\F_2)), \quad x \longmapsto (e_x:f \mapsto f(x)).\]
  Obviously \(e_x\) is a morphism of boolean algebras.
  To show that \(e\) is continuous, let \(f\in \Cont(X,\F_2)\) and \(D(f)\) the associated standard clopen set in \(\Spec(A)\).
  Its preimage under \(e\) is the set of all \(x\in X\) such that \(f(x)=e_x(f)=1\).
  This condition is equivalent to \(f(x)\neq 0\) which is an open condition, so \(e\) is continuous.
  Missing: Bijectivity, naturality.

  Next, consider the evaluation map
  \[e:A \longrightarrow  \Cont(\Spec A,\F_2),\quad a\mapsto (e_a : \phi\mapsto \phi(a))\]
  which is obviously a ring homomorphism.
  To prove injectivity, suppose that \(a\neq 0\).
  By Zorn's lemma, there is a maximal ideal \(I\) of \(A\) which contains \(1-a\neq 1\).
  Since quotient of boolean rings are boolean, and quotients of rings by maximal ideals are fields, the quotient ring \(A/I\) is (uniquely) isomorphic to \(\F_2\).
  By composing this isomorphism with the canonical projection \(A\to A/I\) we obtain a morphism of boolean algebras \(\phi:A\to \F_2\) with kernel \(I\).
  In particular, \(\phi(1-a)=0\), or equivalently, \(\phi(a)=1\), so \(e_a\neq 0\) which proves injectivity.

  For surjectivity, let \(f:\Spec A \to \F_2\) be any continuous function.
  Such a function is uniquely determined by the clopen set \(U=f^{-1}(1)\subseteq\Spec(A)\).
  If we can find an \(a\in A\) with \(U=D(a)\), then \(f\) is exactly \(e_a\).
  Since \(U\) is open, it can be written as a union of standard open sets.
  But since \(U\) is closed and thus compact, finitely many of these suffice, say, \(U=D(a_1)\cup\dots\cup D(a_n)\).
  Recall from above that in a boolean algebra, \(\Spec(A)\setminus D(a)=D(1-a)\).
  Moreover, the standard opens satisfy the property \(D(a)\cap D(b)=D(ab)\) (this holds in the spectrum of every ring).
  Therefore, if we define \(a\) as \(1-(1-a_1)\cdots (1-a_n)\), we obtain
  \[
    D(a) = \Spec(A)\setminus (D(1-a_1)\cap\dots\cap D(1-a_n)) = D(a_1)\cup\dots\cup D(a_n) = U.
  \]
  [The proof of surjectivity corresponds to Exercise 23 (iii) in Atiyah-MacDonald.]
\end{proof}
}

Consider the following three categories.

1) {\em Profinite sets} {\bf ProFin} \\
Objects: $\lim_{i\in I}S_i$ for $S_i$ finite sets, $I$ cofiltered poset
($\forall i,j \exists k\leq i,j$)
\\
Morphisms: 
$$
  \Hom(\lim_{i\in I}S_i,\lim_{j\in J}T_j) := \lim_j\colim_i\Map(S_i,T_j)
$$

2) {\em Totally disconnected compact Hausdorff spaces} {\bf DiscComp}

3) {\em Boolean algebras} {\bf Boolean} \\ 
Objects: commutative rings $A$ such that $x^2=x$ for all $x$ 

\begin{prop}[Stone Duality]
There are equivalences of categories
$$
  \textbf{ProFin} \stackrel{\sim}\longrightarrow 
  \textbf{DiscComp} \stackrel{\sim}\longrightarrow  
  \textbf{Boolean}^{\rm op}.
$$
\end{prop}

\begin{proof}
For 1) $\to$ 2) $\to$ 3), equip $S=\lim S_i$ with the (inverse,
projective) limit topology and assign to it the Boolean algebra
$$
   \Cont(S,\F_2) = \colim_i\Map(S_i,\F_2).
$$
Conversely, assign to $A$ the profinite set
$$
  \Spec A = \Hom(A,\F_2) = \lim_{A_i\subset A\text{ finite}}\Hom(A_i,\F_2).
$$
\end{proof}

There are two measures how ``big'' a profinite set is.

\begin{defi}
Let $S$ be a profinite set. 
\begin{itemize}
\item The {\em size} of $S$ is $\kappa=|S|$.
\item The {\em weight} of $S$ is $\lambda=|\Cont(S,\F_2)|$.
\item $S$ is {\em light} if $\lambda\leq\om=|\N|$.
\end{itemize}
\end{defi}

\begin{rem}
If $\lambda$ is infinite, then $\lambda$ is the smallest
  cardinality of an index set $I$ such that $S=\lim_{i\in I}S_i$.
Thus, $S$ is light iff is is a countable limit of finite sets.
\end{rem}

\begin{example}[Examples of profinite sets]
(0) Finite sets.
  
(1) $S = \N\cup\{\infty\} = \lim_n\{0,1,\dots,n,\infty\}$ has $\kappa=\lambda=\om$. 

(2) The {\em Cantor set} $S = \{0,1\}^\N = \lim_n\{0,1\}^n$ has
  $\kappa=2^\om$ and $\lambda=\om$.

(3) The {\em Stone-\v{C}ech compactification} $S=\beta\N$ satisfies 
$$
  \Cont(\beta\N,\F_2) = \Cont(\N,\F_2) \cong \{\text{subset of }\N\},
$$
so it has $\kappa=2^{2^\om}$ and $\lambda=2^\om$. 
\end{example}

\begin{prop}
(a) We always have $\lambda\leq 2^\kappa$ and $\kappa\leq 2^\lambda$. \\
(b) If $\kappa=\om$ then $\lambda=\om$. 
\end{prop}

\begin{proof}
(a) follows from
\begin{align*}
  \lambda &= |\Cont(S,\F_2)| \leq |\Map(S,\F_2)| = 2^\kappa, \cr
  \kappa &= |\Hom(A,\F_2)| \leq |\Map(A,\F_2)| = 2^\lambda.
\end{align*}
(b) Write $S=\{s_0,s_1,s_2,\dots\}$ and choose quotients
$S\twoheadrightarrow S_n$ which restrict to injections
$\{s_0,\dots,s_n\}\into S_n$. Then the induced map $S\to\lim_n S_n$ is
injective and surjective.
%and thus $\Cont(S,\F_2)=\colim_n\Cont(S_n,\F_2)$ is countable.   
\end{proof}

\begin{rem}
The same proof shows: if $\kappa$ is infinite then $\lambda\leq\kappa$.
\end{rem}

\begin{prop}
The following categories are equivalent to the category of light
profinite sets:

1) $\text{\bf Pro}_{\N}\text{\bf Fin}$ \\
Objects: $\lim_{n\in\N}S_n$ for $S_n$ finite sets
\\
Morphisms: 
\begin{align*}
  \Hom(\lim_{n\in\N}S_n,\lim_{m\in\N}T_m)
  &:= \lim_m\colim_n\Map(S_n,T_m) \cr
  &= \colim_f\lim_m\Map(S_{f(m)},T_m)
\end{align*}
over strictly increasing maps $f:\N\to\N$.
  
2) {\em Metrizable} totally disconnected compact Hausdorff spaces

3) $(\text{{\em Countable} Boolean algebras})^{\rm op}$
\end{prop}

\begin{prop}
The category of light profinite sets has all countable
limits. Sequential limits of surjections are surjective. 
\end{prop}

\begin{prop}
For each light profinite set $S$ there exists a surjection
$\{0,1\}^\N\to S$. 
\end{prop}

\begin{proof}
Exercise (simple induction).
\end{proof}

The following two propositions are special to {\em light} profinite
sets. 

\begin{prop}
Let $S$ be a {\em light} profinite set. Then every open subset $U\subset S$
is a countable union of light profinite sets. 
\end{prop}

\begin{proof}
Write $S=\lim_{n\in\N}S_n$ with $S_n$ finite and the canonical maps
$\pi_n:S\to S_n$. Denote $Z:=S\setminus U$ and $Z_n:=\pi_n(Z)\subset S_n$.
Then $U=\bigcup_n\pi_n^{-1}(S_n\setminus Z_n)$, and thus $Z=\lim_nZ_n$. 
Each $\pi_n^{-1}(S_n\setminus Z_n)\subset S$ is {\em clopen} (closed
and open), and
$$
  U = \coprod_n\Bigl(\pi_{n+1}^{-1}(S_{n+1}\setminus Z_{n+1})\setminus
  \pi_n^{-1}(S_n\setminus Z_n)\Bigr).
$$
\end{proof}

\begin{rem}
There exist open subsets $U\subset S$ of profinite sets $S$ with
sheaf cohomology $H^i(U;\Z)\neq 0$ for some $i>0$, so $U$ is not 
a countable union of profinite sets.
\end{rem}

\begin{prop}
Let $S$ be a nonempty {\em light} profinite set. Then $S$ is an
injective object in {\bf ProFin}, i.e., for each injection $Z\subset
X$ of profinite sets each morphism $Z\to S$ extends to $X$: 
$$
\xymatrix{
  X \ar@{-->}[rd] \\ 
  Z \ar@{^{(}->}[u] \ar[r] & S
}
$$
\end{prop}

\begin{proof}
In the special case $S=\F_2$ the restriction map
$\Cont(X,\F_2)\to\Cont(Z,\F_2)$ is surjective because each clopen
subset of $Z$ extends to a clopen subset of $X$. 
In the general case, write $S=\lim_n S_n$. We may assume that each map
$S_{n+1}\to S_n$ is surjective. By induction over $n$, suppose we
already have extended the map $Z\to S_n$ to $X\to S_n$. By
decomposition into fibers over $S_n$, it suffices to consider the case
$S_n=\{*\}$. Then $S_{n+1}$ is finite and the extension follows
from the special case. 
\end{proof}

{\luca
	\paragraph{Grothendieck Topologies}
	
	Let $\C$ be a (small) category.
	
	\begin{definition}
		A \emph{sieve} S on $U \in \C$ is a family of morphisms with codomain $U$, such that
		\begin{equation*}
			f \in S \quad \Longrightarrow \quad f \circ g \in S
		\end{equation*}
		for all $g$ composable with $f$.
		We denote the set of all sieves on $U$ by $\Omega(U)$.
	\end{definition}
	
	Equivalently a sieve on $U$ can be described as a subobject $S \subseteq h_U \coloneq \Hom_\C(\cdot,U)$ in the category of presheafs $\widehat\C = \{\text{functors : } \C^\op \rightarrow \Set\}$.
	
	For a sieve $S$ on $U$ and a morphism $f : V \rightarrow U$ the restriction along $f$
	\begin{equation*}
		f^*S \coloneq \{g \in h_V \; | \; f \circ g \in S\}
	\end{equation*}
	clearly is a sieve on $V$.
	Using this one can check that $\Omega$ defines a presheaf by setting $\Omega(f) = f^*$.
	
	\begin{definition}
		A \emph{Grothendieck topology} $J$ on $\C$ assigns to each object $U$ a set $J(U) \subset \Omega(U)$ of so called \emph{covering sieves}, such that
		\begin{enumerate}
			\item the maximal sieves $h_U$ is in $J(U)$;
			\item (stability) if $S \in J(U)$, then $f^*S \in J(V)$ for all $f : V \rightarrow U$;
			\item (transitivity) if $S \in J(U)$ and $R \in \Omega(U)$ such that $f^*R \in J(V)$ for all $f : V \rightarrow U$ in $S$, then $R \in J(U)$.
		\end{enumerate}
		The pair $(\C,J)$ is then called a \emph{site}.
	\end{definition}
	
	Because of stability, $J$ is a subobject of $\Omega$ in $\widehat{\C}$.
	
	Here we want to define a Grothendieck topology generated by ``covering families'' that are no sieves in general. For this we assume $\C$ to have pullbacks.
	
	\begin{definition}
		A \emph{basis} $K$ on $\C$ assigns to each object $U$ a set $K(U)$ of families of morphisms with codomain $U$, so called \emph{covers}, such that
		\begin{enumerate}
			\item every isomorphism $\{f : V \rightarrow U\}$ is in $K(U)$;
			\item (stability) if $\{f_i : U_i \rightarrow U\} \in K(U)$, then for all $g : V \rightarrow U$ the family of pullbacks along $g$ $\{g^*f_i : U_i \times_U V \rightarrow V\}$
			is in $K(V)$.
			\item (transitivity) if $\{f_i : U_i \rightarrow U\} \in K(U)$ and for each $i$ one has a cover $\{f_{ij} : U_{ij} \rightarrow U_i\} \in K(U_i)$, then the ``refinement'' $\{f_i \circ f_{ij} : U_{ij} \rightarrow U\}$ is in $K(U)$.
		\end{enumerate}
	\end{definition}
	
	\begin{prop}
		A basis $K$ on $\C$ generates a Grothendieck topology $J$ defined by
		\begin{equation*}
			S \in J(U) \quad : \Longleftrightarrow \quad \exists R \in K(U) : R \subseteq S.
		\end{equation*}
	\end{prop}
	
	\begin{proof}
		\begin{enumerate}
			\item The maximal sieve $h_U$ is in $J(U)$, since it contains the isomorphism $\{\id_U\} \in K(U)$.
			\item Let $S \in J(U)$ with $R \subseteq S$ for some $R \in K(U)$ and $g : V \rightarrow U$ a given morphism.
			Denote by $T \in K(V)$ the pullback of $R$ along $g$ as described in the definition above,
			i.e. for $p \in T$ there is $f \in R$ with a pullback diagram
			$$
			\xymatrix{
				V \times_U W \ar[d]^-p \ar[r]^-q & W \ar[d]^f \\
				V \ar[r]^g & U.
			}
			$$
			Then, because $S$ is a sieve, $g \circ p = f \circ q \in S$.
			Hence $T \subseteq g^*S$, which shows that $g^*S$ is in $J(U)$.
			\item Let $S \in J(U)$ and $R \in \Omega(U)$ such that $f^*R \in J(V)$ for all $f : V \rightarrow U$ in $S$.
			Then we find covers $T \in K(U)$ and $T_f \in K(V)$ for all $f : V \rightarrow U$ in $T$ such that $T \subseteq S$ and $T_f \subseteq f^*R$.
			Now the refinement as described above is a cover contained in $R$.
			Thus $R \in J(U)$.
		\end{enumerate}
	\end{proof}
	
	The category $\Profin$ has pullbacks, which can be seen using Lemma \ref{lem:stone-spaces} in the equivalent category $\Disccomp$.
	The same holds for $\Lightprofin$.
	
	\begin{prop}
		On $\Lightprofin$ finite families of jointly surjective maps ($\{f_i : U_i \rightarrow U\}_{i = 1,\dots,n}$  with $f = (f_i) : \coprod_{i=1}^n U_i \rightarrow U$ surjective) form a basis.
	\end{prop}
	
	\begin{proof}
		\begin{enumerate}
			\item Isomorphisms are surjective.
			\item In $\Lightprofin$ let $\{f_i : U_i \rightarrow U\}_{i = 1,\dots,n}$ be jointly surjective and $g : V \rightarrow U$ given.
			Consider the pullbacks
			$$
			\xymatrix{
				V \times_U U_i \ar[d]^-{p_i} \ar[r]^-{q_i} & U_i \ar[d]^{f_i} \\
				V \ar[r]^g & U.
			}
			$$
			For an arbitrary $v \in V$ there exist by assumption $1 \leq i \leq n$ and $u \in U_i$ such that $f_i(u) = g(v)$.
			Then (by the universal property) $v$ has a preimage under $p_i$.
			Hence $\{p_i : V \times_U U_i \rightarrow V\}_{i = 1,\dots,n}$ is again jointly surjective.
			\item Transitivity is clear since compositions of surjective maps are surjective.
		\end{enumerate}
	\end{proof}
	
	\begin{remark}
		Clearly this is the smallest basis containing the collection of all surjective maps and finite disjoint unions.
		The latter only satisfies (1) and (2) of the basis properties.
	\end{remark}
}

{\luca
	
	\paragraph{Sheafs}
	
	Let $(\C,J)$ be a site.
	
	\begin{definition}
		Let $X : \C^\op \rightarrow \Set$ be a presheaf, $U \in \C$ and $S \in J(U)$ a covering sieve.
		A \emph{matching family} for $S$ of elements of $X$ assigns to each $f : V \rightarrow U$ in $S$ a section $x_f \in X(V)$ in such a way that
		\begin{equation*}
			X(g)(x_f) = x_{f \circ g} \qquad \text{for all } g : W \rightarrow V
		\end{equation*}
		or equivalently: $S \rightarrow X, f \mapsto x_f$ is a natural transformation.
		An \emph{amalgamation} of such a matching family is an element $x \in X(U)$ with
		\begin{equation*}
			X(f)(x) = x_f \qquad \text{for all } f \in S.
		\end{equation*}
		This corresponds to an extension of the natural transformation:
		$$
		\xymatrix{
			S \ar@{^{(}->}[d] \ar[r] & X \\
			h_U \ar@{-->}[ur] &.
		}
		$$
		
		Assume further $\C$ has a basis $K$ and let $R = \{f_i : U_i \rightarrow U\} \in K(U)$ be a cover.
		A matching family for $R$ of elements of $X$ assigns to each $i$ a section $x_i \in X(U_i)$ in such a way that
		\begin{equation*}
			X(\pi_{ij}^1)(x_i) = X(\pi_{ij}^2)(x_j)
		\end{equation*}
		for every pullback square
		$$
		\xymatrix{
			U_i \times_U U_j \ar[d]^-{\pi_{ij}^1} \ar[r]^-{\pi_{ij}^2} & U_j \ar[d]^{f_j} \\
			U_i \ar[r]^{f_i} & U.
		}
		$$
		An associated amalgamation is again an element $x \in X(U)$ with
		\begin{equation*}
			X(f_i)(x) = x_i \qquad \text{for all } i.
		\end{equation*}
	\end{definition}
	
	\begin{definition}
		A presheaf $X : \C^\op \rightarrow \Set$ is a \emph{sheaf} (for J), if every matching family for any covering sieve $S$ of any object $U \in \C$ has a unique amalgamation.
		Diagrammatically this means we have an equalizer diagram
		\begin{equation*}
			\xymatrix{
				X(U) \ar[r]^-e & \prod\limits_{f \in S} X(\dom(f)) \ar@<0.5ex>[r]^-p \ar@<-0.5ex>[r]_-a & \prod\limits_{\substack{f \in S, \\ g \text{ composable}}} X(\dom(g)),
			}
		\end{equation*}
		where the maps are given by
		\begin{align*}
			e(x) & \coloneq (X(f)(x))_f, \\
			p((x_f)_f) & \coloneq (x_{f \circ g})_{f,g}, \\
			a((x_f)_f) & \coloneq (X(g)(x_f))_{f,g}.
		\end{align*}
	\end{definition}
	
	\begin{prop}
		Assume $J$ is generated by a basis $K$.
		Then a presheaf $X : \C^\op \rightarrow \Set$ is a sheaf, iff every matching family for any cover $\{f_i : U_i \rightarrow U\}$ of any object $U \in \C$ has a unique amalgamation.
		Diagrammatically this means we have an equalizer diagram
		\begin{equation*}
			\xymatrix{
				X(U) \ar[r]^-e & \prod\limits_i X(U_i) \ar@<0.5ex>[r]^-{p_1} \ar@<-0.5ex>[r]_-{p_2} & \prod\limits_{i,j} X(U_i \times_U U_j),
			}
		\end{equation*}
		where the maps are given by
		\begin{align*}
			e(x) & \coloneq (X(f_i)(x))_i, \\
			p_1((x_i)_i) & \coloneq (X(\pi_{ij}^1)(x_i))_{i,j}, \\
			p_2((x_i)_i) & \coloneq (X(\pi_{ij}^2)(x_j))_{i,j}.
		\end{align*}
	\end{prop}
	
	\begin{remark}
		Instead of a basis it suffices to use a generator of the topology satisfying only the stability condition (2).
	\end{remark}
}

%%%
\subsection{Light condensed sets}
%%%

\begin{defi}
A {\em light condensed set} is a sheaf on the category of light
profinite sets with the Grothendieck topology generated by
\begin{enumerate}
\item finite disjoint unions;
\item {\em all} surjective maps.
\end{enumerate}
Equivalently, this is a functor 
$$
  X: \text{Pro}_\N\text{Fin}^{\rm op} \to \text{Sets},\qquad S\mapsto X(S)
$$
satisfying
\begin{enumerate}
\setcounter{enumi}{-1}
\item $X(\varnothing)=*$;
\item $X(S_1\amalg S_2)\stackrel{\sim}\longrightarrow X(S_1)\times X(S_2)$;
\item Each surjection $f:T\twoheadrightarrow S$ induces an isomorphism
$$
  f^*:X(S) \stackrel{\cong}\longrightarrow {\rm eq}\Bigl(X(T)
  \overset{p_1^*}{\underset{p_2^*}{\rightrightarrows}} X(T\times_ST)\Bigr).
$$
\end{enumerate}
\end{defi}

{\luca

\begin{proof}
	These are just the sheaf conditions for a minimal collection of generating families.
	(1) generalizes easily to arbitrary finite disjoint unions.
\end{proof}
}

\begin{example}[Key example]
Each topological space $A$ defines a light condensed set 
$$
  \ul{A}: S\mapsto\Cont(S,A).
$$
Here condition (2) is equivalent to the fact that surjections between
compact Hausdorff spaces are quotient maps. We have
\begin{itemize}
\item $\ul{A}(*) = A$ as a set; 
\item $\ul{A}(\N\cup\{\infty\}) = \{\text{convergent sequences in $A$
  (with chosen limit point)}\}$;
\item $\ul{A}(\text{Cantor set})$ carries an action of
  $\End(\text{Cantor set})$. 
\end{itemize}
\end{example}

For any light condensed set $X$, we view $X(\N\cup\{\infty\})$ as the
``convergent sequences in $X$''. 

\begin{remark}[Useless remark]
A light condensed set $X$ is completely determined by $X(\text{Cantor
  set})$ with its action of $\End(\text{Cantor set})$.
\end{remark}

\begin{prop}
The functor
$$
  \text{Top} \to \text{CondSet}^{\rm light},\qquad A\mapsto\ul{A}
$$
has a left adjoint $X\mapsto X(*)_{\rm top}$. Here $X(*)_{\rm top}$ is
the set $X(*)$ equipped with the quotient topology from
$\coprod_{\alpha\in X(\text{Cantor set})}\text{Cantor set} \to X(*)$.
\end{prop}
The map in the proposition is given by regarding $p\in C$ as a map $p\colon \{\ast\} \to C$ and then sending $(p,\alpha)$ for $\alpha\in X(C)$ to $X(p)(\alpha)\in X(\{\ast\})$.

\begin{remark}
$X(*)_{\rm top}$ is a {\em metrizably compactly generated topological
  space}. For any such space $A$ we have an isomorphism
  $\ul{A}(*)_{\rm top}\stackrel{\sim}\to A$. 
\end{remark}

\begin{cor}
The functor
$$
  \{\text{metrizably compactly generated topological space}s\} \into
  \text{CondSet}^{\rm light}
$$
is full and faithful. 
\end{cor}

\begin{rem}[Historical precursors]
1) Johnstone's ``topological topos'', based on just $\N\cup\{\infty\}$
and the canonical topology. This is not finitary and shares the bad
algebraic properties of topoi. 

2) Escordo and Xu considered light profinite sets with only finite
disjoint unions as covers. 
\end{rem}

%%%
\subsection{Light condensed abelian groups}
%%%

Recall that for sheaves on any site, sheaves of abelian groups form a
{\em Grothendieck abelian category}: all limits and colimits exist,
and filtered colimits are exact. 
Thus, $\text{CondAb}^{\rm light}$ is a Grothendieck abelian category. 

\begin{example}
What are the cokernels of the inclusions
$$
  \Q_\disc\into\R \quad\text{and}\quad \R_\disc\to \R?
$$
In the first case we have $\ul{\R}/\ul{\Q_\disc}(*)=\R/\Q$ and 
$$
  (\ul{\R}/\ul{\Q_\disc})(S) = \Cont(S,\R)/\Cont(S,\Q_\disc),
$$
where $\Cont(S,\Q_\disc)$ consists of locally constant maps.
In the second case we have $\ul{\R}/\ul{\R_\disc}(*)=\R/\R=0$ and 
$$
  (\ul{\R}/\ul{\R_\disc})(S) = \Cont(S,\R)/\Cont(S,\R_\disc),
$$
where $\Cont(S,\R_\disc)$ again consists of locally constant maps.
The abelian group $(\ul{\R}/\ul{\R_\disc})(S)$ is e.g.~nontrivial for
$S$ the Cantor set.
\end{example}

\begin{thm}
$\text{CondAb}^{\rm light}$ is a Grothendieck abelian category with
  tensor products and the following properties:
\begin{enumerate}
\item Countable products are exact and satisfy Grothendieck's axiom
  (A86).
\item The free light condensed abelian group $\Z[\N\cup\{\infty\}]$ is
  {\em internally projective}. 
\end{enumerate}
\end{thm}

Here the functor 
$$
  \text{CondAb}^{\rm light} \to \text{CondSet}^{\rm light}
$$
has a left adjoint $X\mapsto \Z[X]$. 
  

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Lecture 3}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%
\subsection{Light condensed sets}
%%%

Why allow the Cantor set $C$?
In any topos there exist intrinsic notions of ``compact'' and
``Hausdorff''. 

\begin{definition}
(1) An object $X$ is {\em quasicompact (qc)} if any cover $\amalg_iX_i\to X$
  admits a finite subcover.
  
(2) An object $X$ is {\em quasiseparated (qs)} if for all quasicompact
$Y,Z\to X$ the fiber product $Y\times_XZ$ is quasicompact. 
\end{definition}

\begin{remark}
For $X\neq\varnothing$ in $\CondSet^\light$, 1) is equivalent to the
existence of a surjection $C\doublearrow{}X$ from the Cantor set, and
2) is equivalent to quasicompactness of $C\times_XC$ for all $f,g:C\to X$. 
If in the definition we only allowed $\N\cup\{\infty\}$ as test
objects, then quasicompact objects would all be countable.
\end{remark}

\begin{prop}
$$
  \{\text{qcqs light condensed sets}\} \cong \{\text{metrizable compact
  Hausdorff spaces}\}
$$
$$
  \{\text{qs light condensed sets}\} \cong {\rm Ind}_{\rm inj}\{\text{metrizable compact
  Hausdorff spaces}\}
$$
\end{prop}

Here for the first equivalence we need that the Grothendieck topology
is finitary and we need to allow the Cantor set. Implicit in it is the
existence of the surjection given by binary expansion
$$
  C=\{0,1\}^\N\to[0,1],\qquad (a_0,a_1,\dots)\mapsto 0.a_1a_2\cdots.
$$
In the second equivalence, the category ${\rm Ind}_{\rm inj}\{\text{metrizable compact
  Hausdorff spaces}\}$ has as objects functors $\{\text{filtered posets}\}\to\{\text{metrizable compact
  Hausdorff spaces}\}$ in which all maps are injections.
  
%%%
\subsection{Light condensed abelian groups}
%%%

Recall that
$$
  \CondAb^\light = \text{abelian group objects in }\CondSet^\light =
         {\rm Shv}(Pro_\N(Fin),Ab).
$$
By general theory of sheaves on topoi, $\CondAb^\light$
\begin{itemize}
\item is a Grothendieck abelian category (i.e., filtered colimits are exact);
\item has a tensor product $M\otimes N=\text{sheafification of
}S\mapsto M(S)\otimes N(S)$;
\item has a unit $\ul{\Z}:S\mapsto \Cont(S,\Z) = \{\text{locally
  constant maps }S\to\Z\}$.
\end{itemize}
Moreover, the functor
$$
   \CondAb^\light \to \CondSet^\light
$$
has a left adjoint functor
$$
   X\mapsto \Z[X].
$$
Here the {\em free condensed abelian group} $\Z[X]$ is the
sheafification of $S\mapsto \Z[X(S)]$. It can be thought of $\Z[X(*)]$
equipped with some ``topology''. 

\begin{example}
How to define a topology on
$$
  \Z[\R] = \{\sum_{x\in\R}n_x[x]\mid n_x\in\Z\text{ almost all }0\}\,?
$$
View it as the free condensed abelian group
$$
  \Z[\ul{\R}] = \colim_{I=[-c,c]\subset\R}\Z[I],\qquad
  \Z[I]=\cup_{n\in\N}\Z[I]_{\leq n} 
$$
with the metrizable compact Haudorff spaces (Exercise!)
$$
  \Z[I]_{\leq n} := \{\sum_{x\in I}n_x[x]\mid \sum_x|n_x|\leq n\}.
$$
Note that $\Z[\R]$ is actually a condensed {\em ring}. 
\end{example}

{\bf Notation. }
Here and in the sequel we identify a topological space $A$ with the
associated light condensed set $\ul{A}$ (``Yoneda embedding'') and drop
the underscore. 

\begin{thm}
In $\CondAb^\light$ the following hold:
\begin{enumerate}
\item Countable products are exact.
\item Sequential limits of surjective maps are surjective.
\item $\Z[\N\cup\{\infty\}]$ is internally projective.
\end{enumerate}
\end{thm}

Here an object $P$ is called
\begin{itemize}
\item {\em projective} if $\Ext^i(P,-)=0$ for all $i>0$;
\item {\em internally projective} if $\ul{\Ext}^i(P,-)=0$ for all $i>0$
\end{itemize}
where $\ul{\Ext}^i(M,N)$ is the sheafification of
$S\mapsto\Ext^i(M\otimes\Z[S],N)$. 

\begin{proof}
\ul{Reduction of (1) to (2)}: 
All properties are clear except for surjectivity. For this we need
that for surjections $f_n:M_n\doublearrow{}N_n$ the product
$$
  \prod f_n:\prod M_n\to \prod N_n
$$
is also surjective. This follows by writing it as the limit of surjections
$$
  \lim_m\Bigl(\prod_{n\leq m}M_m\times\prod_{n>m}N_n\doublearrow{}\prod_nN_n\Bigr).
$$
(2) We need to prove surjectivity of
$$
  M_\infty := \lim\Bigl(\cdots
  M_2\doublearrow{}M_1\doublearrow{}M_0\Bigr)\to M_0. 
$$
This means that for each $S\in Pro_\N(Fin)$ and $S\to M_0$ we need to find $S_\infty$
such that 
$$
\xymatrix{
  M_\infty \ar[r] & M_0 \\
  S_\infty \ar[u] \ar@{->>}[r] & S \ar[u].
}
$$
We inductively find $S_n\to M_n$ such that
$$
\xymatrix{
  \cdots M_2 \ar@{->>}[r] & M_1 \ar@{->>}[r] & M_0 \\
  \cdots S_2 \ar[u] \ar@{->>}[r] & S_1 \ar[u] \ar@{->>}[r] & S \ar[u]
}
$$
and set $S_\infty := \lim\Bigl(\cdots S_2\doublearrow{}S_1\doublearrow{}S\Bigr)$.
The map $S_\infty\to S$ is surjective because countable limits of
surjections are surjective in $Pro_\N(Fin)$. 

(3) Note that $\Z[\infty$ is a direct factor in $\Z[\N\cup\{\infty\}]$
  and $M:=\Z[\N\cup\{\infty\}]/\Z[\infty]$ classifies null sequences.
It suffices to show that $M$ is internally projective. For
projectivity, for a surjection of light abeliian groups $\wt N\to N$
and $f:M\to N$ we need to find a lift
$$
\xymatrix{
  & \wt N \ar@{->>}[d] \\ M \ar[ur]^{?} \ar[r]^f & N
}
$$
By surjectivity we find $S\in Pro_\N(Fin)$ such that
$$
\xymatrix{
  S \ar@{->>}[d] \ar[rr]^g & & \wt N \ar@{->>}[d] \\ \N\cup\{\infty\} \ar[r] & M \ar[r]^f & N
}
$$
where $\N\cup\{\infty\}\to N$ sends $\infty\mapsto 0$. 
Picking in $S$ a preimage of each $n\in\N$ together with the full
preimage of $\infty$ gives a closed subset $S'\subset S$ such that
$S'\times_{\N\cup\{\infty\}}\N\stackrel{\sim}\to\N$. After replacing $S$
by $S'$ we may assume w.l.o.g.~that $S\times_{\N\cup\{\infty\}}\N\stackrel{\sim}\to\N$.
Let
$$
  S_\infty := S\times_{\N\cup\{\infty\}}\{\infty\}\subset_{closed}S.
$$
Since $S_\infty$ is light, it is injective in $Pro(Fin)$, so there
exists a retraction $r:S\to S_\infty$. Denote by $i:S_\infty\into S$
the inclusion. Since
$$
\xymatrix{
  S_\infty \ar[d] \ar@{^{(}->}[r] & S \ar@{->>}[d] \\
  \{\infty\} \ar@{^{(}->}[r] & \N\cup\{\infty\}
}
$$
is a pushout in light condensed sets, we get a diagram
$$
\xymatrix{
  S \ar[dr] \ar[rr]^{g-g\circ i\circ r} & & \wt N \\
  & \N\cup\{\infty\} \ar[ur]^{\wt f} & \\
  & \infty \ar[dr] \ar@{^{(}->}[u] & \\
  S_\infty \ar@{^{(}->}[uuu] \ar[ur] \ar@{->>}[rr] & & 0 \ar@{^{(}->}[uuu]
}
$$
It remains to show that the following diagram commutes:
$$
\xymatrix{
  M = \Z[\N\cup\{\infty\}]/\Z[\infty] \ar[dr]_f
  \ar[r]^{\ \ \ \ \ \ \ \ \ \ \ \ \wt f} & \wt N \ar[d] \\
  & N
}
$$
It suffices to check this for the composition with the map $\Z[S]\to
M$, which follows from the fact that $g\circ i\circ r$ projects to $0$
in $N$.
This proves projectivity, and internal projectivity follows by
essentially the same argument. 
\end{proof}

\begin{remark}
For (2) it is crucial that countable limits of covers are covers in
the Grothendieck topology. This forces one to use totally disconnected
spaces as building blocks. For example, a sequence of covers of an
open interval by more and more open intervals converges in teh limit
to a totally disconnected space. 

Part (3) is only true in the light setting. 
\end{remark}

\begin{remark}[Comparison to $\CondAb$]
In $\CondAb$ all products are exact and we have projective generators
$\Z[S]$, where $S=\beta I$ is the Stone-\v{C}ech compactification
of a discrete set $I$. But none of these generators are internally
projective because $\Z[\beta I\times\beta J]$ is never projective for
infinite $I,J$. Similarly, $\Z[\N\cup\{\infty\}]$ is not projective in
$\CondAb$ because $\Z[\beta\N]\doublearrow{}\Z[\N\cup\{\infty\}]$ does
not split (Exercise). 
\end{remark}

\begin{remark}[Relation to Banach spaces]
(surjectivity becomes injectivity).
\begin{itemize}
\item The only known injective Banach spaces are $\Cont(S,\R)$ for $S$
  {\em extremely disconnected}, i.e., a retract of $\beta I$ above.  
\item The space of null sequences $c_0(\N)$ is not injective, but it
  is ``separably injective'', i.e., injective when tested on separable
  Banach spaces. 
\end{itemize}
\end{remark}


%%%
\subsection{Cohomology}
%%%

The following is a special case of the \'etale cohomology on a topos. 

\begin{definition}
For a light condensed set $X$ and an abelian group $M$ we define
$$
  H^i(X,M) := \Ext^i_{\CondAb^\light}(\Z[X],\ul{M}).
$$
\end{definition}

\begin{thm}
If $X$ is a CW complex, then
$$
   H^i(\ul{X},M) \cong H^i_{\rm sing}(X,M).
$$
\end{thm}

The proof is based on the Dold-Thom theorem, according to which
$\Z[\ul{X}]$ is a model for the homology of $X$. 


\printbibliography

\end{document} 


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% JUNK %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%




